Reasonably Programmable Literal Notation

### 2018-06-25 icfp18/syntax-icfp18.tex
Cleaned up bib
added brief mention of recent work by Pombrio and Krishnamurthy
added brief mention of Rust macros

Some programming languages, notably including theorem provers like Agda \cite{norell2007towards} and Coq \cite{Coq:manual}, support ``mixfix'' notation directives \cite{wieland2009parsing,missura1997higher,5134}. Many of these systems enforce \textbf{Capture Avoidance} and application-site \textbf{Context Independence} \cite{5134,DBLP:conf/gpce/TahaJ03,Coq:manual,DBLP:conf/ifl/DanielssonN08}. The problem is that mixfix notation requires a fixed number of sub-trees, e.g. \li{if _ then _ else _}. Coq has some \emph{ad hoc} extensions for list-like literals \cite{Coq:manual}. These systems cannot express the example literal forms from this paper, because they can have any number of spliced terms.

@@

\citet{conf/icfp/LorenzenE13,conf/popl/LorenzenE16}'s SoundExt is a grammar-based syntax extension system where extension providers can equip their new forms with derived typing rules. The system then attempts to automatically verify that the expansion logic, expressed using a rewrite system, rather than an arbitrary function, is sound relative to these derived rules, so it is possible to reason about \textbf{Expansion Typing}. SoundExt does not enforce hygiene, i.e. expansions might depend on the context and intentionally induce capture. A client can only indirectly reason about binding by inspecting the derived typing rules. There is no abstract segmentation discipline. Unlike TLMs, SoundExt supports type-dependent expansions \cite{conf/popl/LorenzenE16}. The trade-off is that TLMs can generate proto-expansions, and therefore segmentations, even when the spliced expressions are ill-typed. Another important distinction is that TLMs rely on proto-expansion validation, rather than verification as in SoundExt (though providers can use MetaOCaml and related systems to reason about typing, as described in Sec. \ref{sec:metaocaml}). The trade-off is that TLMs do not require a fully mechanized host language definition. Finally, there is no clear notion of parameterization or partial application in these systems, so it would be difficult to define notation for, for example, all finite map implementations as we demonstrated in Sec.~\ref{sec:ptsms}.

>>>>

Some programming languages, notably including Rust \cite{Rust/Macros} and theorem provers like Agda \cite{norell2007towards} and Coq \cite{Coq:manual}, support ``mixfix'' notation directives \cite{wieland2009parsing,missura1997higher,5134}. Many of these systems enforce \textbf{Capture Avoidance} and application-site \textbf{Context Independence} \cite{5134,DBLP:conf/gpce/TahaJ03,Coq:manual,DBLP:conf/ifl/DanielssonN08}. The problem is that mixfix notation requires a fixed number of sub-trees, e.g. \li{if _ then _ else _}. Rust and Coq have some \emph{ad hoc} extensions for list-like literals \cite{Rust/Macros,Coq:manual}. These systems cannot express the example literal forms from this paper, because they can have any number of spliced terms.

@@

\citet{conf/icfp/LorenzenE13,conf/popl/LorenzenE16}'s SoundExt is a grammar-based syntax extension system where extension providers can equip their new forms with derived typing rules. The system then attempts to automatically verify that the expansion logic, expressed using a rewrite system, rather than an arbitrary function, is sound relative to these derived rules, so it is possible to reason about \textbf{Expansion Typing}. Recent work by \citet{DBLP:conf/pldi/PombrioK18} instead infers the typing rules. These systems do not enforce hygiene, i.e. expansions might depend on the context and intentionally induce capture. A client can only indirectly reason about binding by inspecting the derived typing rules. There is no abstract segmentation discipline. Unlike TLMs, SoundExt supports type-dependent expansions \cite{conf/popl/LorenzenE16}. The trade-off is that TLMs can generate proto-expansions, and therefore segmentations, even when the spliced expressions are ill-typed. Another important distinction is that TLMs rely on proto-expansion validation, rather than verification as in SoundExt (though providers can use MetaOCaml and related systems to reason about typing, as described in Sec. \ref{sec:metaocaml}). The trade-off is that TLMs do not require a fully mechanized host language definition. Finally, there is no clear notion of parameterization or partial application in these systems, so it would be difficult to define notation for, for example, all finite map implementations as we demonstrated in Sec.~\ref{sec:ptsms}.
